A python toolbox to tackle the curse of imbalanced datasets in machine learning

### 2016-06-21 main.tex
Advance the writting

\begin{abstract}
\emph{UnbalancedDataset} is an open-source python toolbox aiming at providing a wide range of methods to cope with the problem of imbalanced dataset frequently encountered in machine learning and pattern recognition.
The state-of-the-art methods implemented can be categorized into 4 different sampling strategies: (i) under-sampling, (ii) over-sampling, (iii) combination of over- and under-sampling, and (iv) ensemble learning methods.
The proposed toolbox only depends of \emph{numpy}, \emph{scipy}, and \emph{scikit-learn} and is distributed under MIT license.
The implementation pattern is identical to the \emph{scikit-learn} API to offer compatibility.
Documentation, unit tests as well as integration tests are provided to allow new developers to contribute.
The toolbox is publicly available in GitHub \url{https://github.com/fmfn/UnbalancedDataset} as well as the documentation \url{http://fmfn.github.io/UnbalancedDataset/}.
\end{abstract}

@@

Many real world datasets have many samples of some classes (majority classes),
and only a few samples the other class (minority classes). This imbalance gives
rise to the ``class imbalance'' problem~\citep{prati2009data} (or ``curse of imbalanced datasets'')
which is the problem of learning a concept from the class that has a small number of samples compared
to the other classes.

The class imbalance problem has been encountered in multiple areas such as 
telecommunication managements, bioinformatics, fraud detection, and medical diagnosis,
and has been considered one of the top 10 problems in data mining and 
pattern recognition~\citep{rastgoo2016tackling,yang200610}. 
Imbalanced data substantially compromises the learning process, since most of the 
standard machine learning algorithms expect balanced class distribution or an 
equal misclassification cost~\citep{he2009learning}. For this reason, several
approaches have been specifically proposed to handle such datasets.

Such standalone methods have been implemented mainly in R language~\citep{torgo2010data,kuhn2015caret,dal2013racing}.
Up to our knowledge, however, there is no python toolbox allowing such processing while cutting edge machine learning toolboxes are available~\citep{pedregosa2011scikit,sonnenburg2010shogun}.

In this paper, we present the \textbf{UnbalancedDataset} API, 
\textit{a python toolbox to tackle the curse of imbalanced datasets
in machine learning}. The following sections present the implemented
methods, implemented design, sustainability and continuous integration details,
and finally, the conclusion of this paper, including future functionalities
for the UnbalancedDataset API.

%This problem is frequently referred as ``class imbalance'' problem~\citep{prati2009data} and has been encountered in multiple areas such as telecommunication managements, bioinformatics, fraud detection, and medical diagnosis. 
%Imbalanced data substantially compromises the learning process since most of the standard machine learning algorithms expect balanced class distribution or an equal mis-classification cost~\citep{he2009learning}.
%Medical data are prone to such drawbacks due to the fact that the portion of diseased samples or patients is far lower than healthy cases.
%Furthermore, the detection and classification of minority malignant cases are highly essential so that the sensitivity of developed algorithms need to be maximized.
%Consequently, the problem of imbalanced data is usually addressed by employing different techniques which do not impair the topology of the data.

@@

\noindent\textit{Documentation} A consistent API documentation is provided using \emph{sphinx} and \emph{numpydoc}.
Additional installation guide, examples, and tutorial are also provided and centralized on GitHub\footnote{\url{http://fmfn.github.io/UnbalancedDataset/}}.

@@

The implementation rely on \emph{numpy}, \emph{scipy}, and \emph{scikit-learn}.
Each class implements 3 main functions inspired from the \emph{scikit-learn} API:
(i) \emph{fit} computes the parameter values which are later needed to transform the data into a balanced set;
(ii) \emph{transform} performs the sampling and return the data with the desired balancing ratio;
and (iii) \emph{fit\_transform} is equivalent of calling the function \emph{fit} follow the function \emph{transform}.

@@

The \textbf{UnbalancedDataset} toolbox provides four different strategies to tackle the problem of imbalanced dataset:
(i) under-sampling, (ii) over-sampling, (iii) a combination of both, and (iv) ensemble learning.
The following sections give an overview of the techniques implemented.

@@

In the contrary, the data balancing can be performed by over-sampling in which the new samples belonging to the minority class are generated aiming at equalizing the number of samples in both classes.
Two different methods are considered.

\textbf{Random over-sampling} is performed by randomly replicating the samples of the minority class such that the number of samples is equal in both minority and majority classes.

@@

\textbf{SMOTE + Tomek links} are combined to clean the samples created using SMOTE~\citep{batista2003balancing}.
SMOTE over-sampling can lead to over-fitting which can be avoided by removing the Tomek links from both majority and minority classes~\citep{prati2009data}.
\textbf{SMOTE + edited nearest neighbours} are combined for the same aforementioned reason~\citep{batista2004study}.

>>>>

\begin{abstract}
\texttt{UnbalancedDataset} is an open-source python toolbox aiming at providing a wide range of methods to cope with the problem of imbalanced dataset frequently encountered in machine learning and pattern recognition.
The state-of-the-art methods implemented can be categorized into 4 different sampling strategies: (i) under-sampling, (ii) over-sampling, (iii) combination of over- and under-sampling, and (iv) ensemble learning methods.
The proposed toolbox only depends of \texttt{numpy}, \texttt{scipy}, and \texttt{scikit-learn} and is distributed under MIT license.
Documentation, unit tests as well as integration tests are provided to ease usage and contribution.
The toolbox is publicly available in GitHub \url{https://github.com/fmfn/UnbalancedDataset}.
\end{abstract}

@@

Real world datasets commonly show the particularity to have the number of samples of a class under-represented compared to other classes.
This imbalance gives rise to the ``class imbalance'' problem~\citep{prati2009data} (or ``curse of imbalanced datasets'')
which is the problem of learning a concept from the class that has a small number of samples compared
to the other classes.

The class imbalance problem has been encountered in multiple areas such as 
telecommunication managements, bioinformatics, fraud detection, and medical diagnosis,
and has been considered one of the top 10 problems in data mining and 
pattern recognition~\citep{rastgoo2016tackling,yang200610}. 
Imbalanced data substantially compromises the learning process, since most of the 
standard machine learning algorithms expect balanced class distribution or an 
equal misclassification cost~\citep{he2009learning}. For this reason, several
approaches have been specifically proposed to handle such datasets.
Such standalone methods have been implemented mainly in R language~\citep{torgo2010data,kuhn2015caret,dal2013racing}.
Up to our knowledge, however, there is no python toolbox allowing such processing while cutting edge machine learning toolboxes are available~\citep{pedregosa2011scikit,sonnenburg2010shogun}.

In this paper, we present the \texttt{UnbalancedDataset} API, 
\textit{a python toolbox to tackle the curse of imbalanced datasets
in machine learning}. The following sections present the implemented
methods, implemented design, sustainability and continuous integration details,
and finally, the conclusion of this paper, including future functionalities
for the \texttt{UnbalancedDataset} API.

@@

\noindent\textit{Documentation} A consistent API documentation is provided using \texttt{sphinx} and \texttt{numpydoc}.
Additional installation guide, examples, and tutorial are also provided and centralized on GitHub\footnote{\url{http://fmfn.github.io/UnbalancedDataset/}}.

@@

The implementation rely on \texttt{numpy}, \texttt{scipy}, and \texttt{scikit-learn}.
Each class implements 3 main functions inspired from the \texttt{scikit-learn} API:
(i) \texttt{fit} computes the parameter values which are later needed to transform the data into a balanced set;
(ii) \texttt{transform} performs the sampling and return the data with the desired balancing ratio;
and (iii) \texttt{fit\_transform} is equivalent of calling the function \texttt{fit} follow the function \texttt{transform}.

@@

The \texttt{UnbalancedDataset} toolbox provides four different strategies to tackle the problem of imbalanced dataset:
(i) under-sampling, (ii) over-sampling, (iii) a combination of both, and (iv) ensemble learning.
The following sections give an overview of the techniques implemented.

@@

In the contrary of under-sampling, data balancing can be performed by over-sampling in which new samples are generated in $\chi_{min}$ to reach the balancing ratio $r_{\chi_{res}}$.
The following methods are currently available.

\textbf{Random over-sampling} is performed by randomly replicating the samples of $\chi_{min}$ to obtain the appropriate balancing ratio $r_{\chi_{res}}$.

@@

\textbf{SMOTE + Tomek links} are combined to clean the samples created using SMOTE~\citep{batista2003balancing}.
SMOTE over-sampling can lead to over-fitting which can be avoided by removing the Tomek links from both majority and minority classes~\citep{prati2009data}.

\textbf{SMOTE + edited nearest neighbours} are combined for the same aforementioned reason~\citep{batista2004study}.
